A Survey of Quantitative Descriptions of Molecular Structure

### 2012-04-23 paper.tex
Added the UML diagram

Due to our experience with the CDK project we provide a brief overview
of the architecture of the descriptor calculations implemented by the
CDK. First, the toolkit explicitly differentiates between atom, bond
and whole molecule descriptors. There are a total of 30, 6 and 51
descriptors for atoms, bonds and molecules respectively. Note that
each individual descriptor may calculate multiple values. For example,
the CPSA descriptor evaluates 30 individual values. Thus for a given
molecule it is possible to generate approximately 300 descriptor
values. Every descriptor implements an interface allowing them to
return values in a uniform manner, viz., via a
\texttt{DescriptorValue} object. This class can represents the result
of a descriptor calculation and includes both the descriptor values
(via a \texttt{IDescriptorResult} object) as well as meta-data on the
descriptor itself via a \texttt{DescriptorSpecification} object (see
below for more details). The descriptor classes are designed such that
in the event of a calculation error (invalid, missing or nonsensical
input), null values are returned, rather than throwing an
exception. This is useful during automated descriptor calculation,
where one desires a ``rectangular'' descriptor matrix, from which one
could filter out the undefined descriptors. While the user can pick
and choose specific descriptors to calculate, the API provides a class
(\texttt{DescriptorEngine}) to specifically evaluate all or a subset
of descriptors for a molecule at one go.  Currently the descriptor
implementations are not threadsafe, making them unsuitable for
parallelization. This is a drawback since many of the descriptors can
be slow. For example, the $\chi$ descriptors \cite{Kier:1976aa} are
implemented using SMARTS matching to identify the various paths and
chains described by Kier \textit{et al.}, rather than explicitly walking the
molecular graph.

@@

>>>>

Due to our experience with the CDK project we provide a brief overview
of the architecture of the descriptor calculations implemented by the
CDK. First, the toolkit explicitly differentiates between atom, bond
and whole molecule descriptors. There are a total of 30, 6 and 51
descriptors for atoms, bonds and molecules respectively. Note that
each individual descriptor may calculate multiple values. For example,
the CPSA descriptor evaluates 30 individual values. Thus for a given
molecule it is possible to generate approximately 300 descriptor
values. Every descriptor implements an interface allowing them to
return values in a uniform manner, viz., via a \texttt{DescriptorValue}
object (see Figure~\ref{fig:cdkapi}). This class can represents the result
of a descriptor calculation and includes both the descriptor values
(via a \texttt{IDescriptorResult} object) as well as meta-data on the
descriptor itself via a \texttt{DescriptorSpecification} object (see
below for more details). The descriptor classes are designed such that
in the event of a calculation error (invalid, missing or nonsensical
input), null values are returned, rather than throwing an
exception. This is useful during automated descriptor calculation,
where one desires a ``rectangular'' descriptor matrix, from which one
could filter out the undefined descriptors. While the user can pick
and choose specific descriptors to calculate, the API provides a class
(\texttt{DescriptorEngine}) to specifically evaluate all or a subset
of descriptors for a molecule at one go.  Currently the descriptor
implementations are not threadsafe, making them unsuitable for
parallelization. This is a drawback since many of the descriptors can
be slow. For example, the $\chi$ descriptors \cite{Kier:1976aa} are
implemented using SMARTS matching to identify the various paths and
chains described by Kier \textit{et al.}, rather than explicitly walking the
molecular graph.

@@

\ctable[caption={UML diagram of the CDK descriptor API in version 1.4.},cap={},label={fig:cdkapi},botcap,figure]
{c}
{}
{
  \includegraphics[width=\linewidth]{iDescriptor.pdf}
}
